Individual heterogeneity in the functional network topography of the medial prefrontal cortex

### 2019-05-01 Manuscript.Rmd
working on discussion

A considerable amount of meta-analytic work has been dedicated to pinpointing the spatial embedding of behavioral constructs in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across psychological phenomena [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, the effects elicited by the subjective value of an option and the default network (DN) have been suggested to be inseparable in this area, an overlap that partially extends to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. However, recent studies have shown considerable variability of the brain's functional topography across individuals, emphasizing the importance of capturing its organization at the individual subject level, and pointing to limitations in aggregative estimations [@Braga2017; @Gordon2017; @Gratton2018b; @Kong2018; @Laumann2015; @Tobyne2018]. Therefore, determining the distinct functional mapping of psychological processes (and the potential extent of their overlap) first requires: 1) understanding and quantifying the degree of heterogeneity in the intrinsic functional organization of this area in a large group of individuals; and 2) a method to reliably and precisely capture it in isolated subjects.

Here we address these challenges by using spectral partitioning, a graph-theoretic clustering algorithm, to subdivide areas of overlap into DN and non-DN in 100 individuals from the Human Connectome Project. Our results show that the intrinsic functional topography of mPFC and PCC can be reliably partitioned into at least two idiosyncratically arranged networks that can be obscured when averaged across subjects. These sub-networks were stable over time and across test-retest days using moderate amounts of fMRI data (up to 30 min per day), and resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. We also show that spectral partitioning is capable of identifying these sub-networks in isolated individuals better than more direct seed-based functional connectivity approaches.

@@

  - It matters whether distinct cognitive functions overlap in the brain. Apparent overlap has motivated theoretical proposals about ways in which seemingly dissimilar cognitive domains might recruit a shared set of core cognitive processes.

  - The interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains. Individual-level effects wash out in group averages (Michalka), apparent overlap in functional activation can vanish with increases in spatial precision (Woo), and resting-state networks have reliable individual-specific characteristics (Muller, Braga).

  - We don't yet really understand how much individual-level heterogeneity there is in the functional topography of mPFC. We need to (1) develop an efficient/robust method to capture individual-level topography, and (2) quantify the degree of heterogeneity in a large group of individuals.

  - The method we propose is spectral partitioning. Restricting our analyses to a general mPFC/PCC search space makes it appropriate to use a technique that identifies a vertex-wise, binary partitioning. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. SP has advantages (deterministic, simple diagnostics). Comparisons with modularity maximization suggested the binary partitioning was appropriate.

>>>>

A considerable amount of meta-analytic work has been dedicated to pinpointing the spatial embedding of behavioral constructs in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across psychological phenomena [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, the effects elicited by the subjective value of an option and the default network (DN) have been suggested to be inseparable in this area, an overlap that partially extends to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. This apparent neural overlap has important implications, as it has motivated theoretical proposals about ways in which these seemingly dissimilar cognitive domains might recruit a shared set of core cognitive processes [@Northoff2012].

However, the interpretation of overlap in group-level data depends on the degree to which functional organization is heterogeneous across individuals. Recent studies have shown that heteromodal brain regions have considerable variability in functional connectivity across individuals [@Muller2013a], individual-specific functional topography can be occluded in aggregative estimations [@Braga2017; @Gordon2017; @Michalka2015; @Tobyne2018], and overlap in functional activation can vanish with increases in spatial precision [@Woo2014]. These findings suggest that group-level and meta-analysis-level overlap does not necessarily imply overlap in individual brains.

So far, our understanding of the individual-level heterogeneity in the functional topography of mPFC has been mostly descriptive. Therefore, affirming the extent of overlap among psychological processes in this area first requires: 1) a method to reliably and precisely capture its functional topography in isolated individuals; and 2) quantifying the degree of heterogeneity of its intrinsic functional organization in a large group of individuals.

Here we address these challenges by using spectral partitioning, a graph-theoretic clustering algorithm, to reliably subdivide canonical DN and limibic networks [@Yeo2011] on the medial cortex into DN and non-DN in 100 individuals from the Human Connectome Project. Restricting our analyses to this general mPFC/PCC search space makes it appropriate to use a technique that identifies a vertex-wise, binary partitioning. This contrasts with whole-brain network analyses, which need to allow for multiple sub-networks and which often use coarser-grained parcels as the units of analysis. SP has advantages (deterministic, simple diagnostics). Comparisons with modularity maximization suggested the binary partitioning was appropriate.

Our results show that the intrinsic functional topography of mPFC and PCC can be reliably partitioned into at least two idiosyncratically arranged networks that can be obscured when averaged across subjects. These sub-networks were stable over time and across test-retest days using moderate amounts of fMRI data (up to 30 min per day), and resembled an organizational scheme recently observed in a handful of subjects [@Braga2017]. We also show that spectral partitioning is capable of identifying these sub-networks in isolated individuals better than more direct seed-based functional connectivity approaches.

@@
